Bandit Multiclass Classification

### 2018-10-12 ptf_margin.tex
.

@@

\section{Construction of polynomials}
Our $\phi$'s in considerations are as follows: TODO: fix the constant in phi so that it has norm at most 1.
\begin{enumerate}
\item $\phi(x) = (x_1^{i_1} \ldots x_d^{i_d})_{i_1 + \ldots + i_d \leq \deg, i_1, \ldots, i_d \geq 0} = (x_S)_{S \in N^d, \|S\|_1 \leq \deg}$. This corresponds to the complete symmetric kernel $K_{\deg}$.
\item $\phi(x) = ( \sqrt{ {\deg \choose i_1+\ldots+i_d} {i_1+\ldots+i_d \choose i_1 \ldots i_d}}  x_1^{i_1} \ldots x_d^{i_d})_{i_1 + \ldots + i_d \leq \deg, i_1, \ldots, i_d \geq 0}$. This corresponds to the polynomial kernel
$K_{\deg}'(x,x') = (1 + \inner{x, x'})^{\deg}$.
\item $\phi(x) = ( \sqrt{ (\frac 1 2)^{i_1+\ldots+i_d} {i_1+\ldots+i_d \choose i_1 \ldots i_d}}  x_1^{i_1} \ldots x_d^{i_d})_{i_1 + \ldots + i_d \leq \deg, i_1, \ldots, i_d \geq 0}$. This corresponds to the kernel $K(x,x') = \frac{1}{1 - \frac 1 2 \inner{x,x'}}$.
\end{enumerate}

@@

\subsection{Construction using rational functions}
\paragraph{The construction.} Define
\[
  Q(x) = \sum_{i=1}^t S_{\ceil{\log \frac 1 \gamma}}^{2t}(\frac{v_i \cdot x}{\gamma}) - (t - \frac 1 2).
\]
where function
\[
S_n^k(x) = \frac{P_n(-x)^{h(k)} - P_n(x)^{h(k)}}{P_n(-x)^{h(k)} + P_n(x)^{h(k)}} = \frac{A_n^k(x)}{B_n^k(x)},
\]
$h(k)$ is the smallest odd integer greater than $\log(2k+1)$, and $P_n(x) = (x - 1) \cdot \prod_{i=1}^n (x - 2^i)^2$.
Denote by $D_i(x) = S_{\ceil{\log \frac 1 \gamma}}^{2t}(\frac{v_i \cdot x}{\gamma})$. $Q$ can therefore be simplified to 
\[ Q(x) = \sum_{i=1}^t D_i(x) - (t - \frac 1 2). \]

\begin{lemma}[~\cite{klivans2004learning}]
For all $(x,y)$ in $S$ satisfying Definition~\ref{def:int-margin}, we have that $y Q(x) \geq \frac 1 2$.
\end{lemma}
\begin{proof}
We consider two cases regarding the label of the example.
\begin{enumerate}
\item Suppose $y = 1$, then for all $i$ in $[t]$, $v_i \cdot x \in [\gamma, 1]$. This implies that $\frac{v_i \cdot x}{\gamma} \in [1, \frac 1 \gamma]$. 
By item~\ref{item:sign-rt} of Fact~\ref{fact:newman}, we have that $D_i(x) \in [1, 1+ \frac 1 {2t}]$. 
Hence, $Q(x) \geq t - (t - \frac 1 2) \geq \frac 1 2$.

\item Suppose $y = -1$. First observe that for all $i$ in $[t]$, $\frac{v_i \cdot x}{\gamma} \in [-\frac 1\gamma, \frac 1 \gamma]$. By items~\ref{item:sign-rt},
~\ref{item:sign-lt} and~\ref{item:sign-mid} of Fact~\ref{fact:newman}, we have that $S_{\ceil{\log \frac 1 \gamma}}^{2t}(\frac{v_i \cdot x}{\gamma}) \leq 1+ \frac 1 {2t}$.
In addition, by Definition~\ref{def:int-margin}, there exists an $i_0$ in $[t]$, such that $v_{i_0} \cdot x \leq -\gamma$. 
By item~\ref{item:bound-out} of Fact~\ref{fact:cheb}, we have that
$D_{i_0}(x) \in [-1 - \frac 1 {2t}, -1]$.
Hence, 
\[ Q(x) = D_{i_0}(x) + \sum_{i \neq i_0} D_i(x) - (t - \frac 1 2)
\leq -1 + (t-1)(1 + \frac 1 {2t}) - (t - \frac 1 2) \leq - \frac 1 2. \]
%\leq -1 + t-1 + \frac{t-1}{2t} - (t - \frac 1 2)

\end{enumerate}
\end{proof}

\begin{fact}
$S_n^k(\cdot)$ has the following properties:
\begin{enumerate}
\item If $x$ is in $[1, 2^n]$, then $S_n^k(x)$ is in $[1, 1 + \frac 1 k]$. \label{item:sign-rt}
\item If $x$ is in $[-2^{-n}, -1]$, then $S_n^k(x)$ is in $[-1 - \frac 1 k, -1]$. \label{item:sign-lt}
\item If $x$ is in $[-1, 1]$, then $|S_n^k(x)|$ is in $[-1, 1]$. \label{item:sign-mid}
\item For all $x$ in $[-2^{-n}, 2^{-n}]$, $\abs{B_n^k(x)} \geq 2^{\frac{n(n-1)h(k)}2 - 1}$. \label{item:sign-denom}
\end{enumerate}
\label{fact:newman}
\end{fact}
\begin{proof}
We prove all the items accordingly.
\begin{enumerate}
\item Observe that $S_n^k(x)$ can be written as:
\[ S_n^k(x) = \frac{1 + (-\frac{P_n(x)}{P_n(-x)})^{h(k)}}{ 1 - (-\frac{P_n(x)}{P_n(-x)})^{h(k)} }.  \]
Denote by $r = (-\frac{P_n(x)}{P_n(-x)})^{h(k)}$. By item~\ref{item:x-large} of Fact~\ref{fact:pn}, $r \in [0, \frac{1}{2k+1}]$. This implies that 
$S_n^k(x) = \frac{1+r}{1-r} \in [1, 1 + \frac 1 k]$.

@@

\item If $x$ is in $[1,2^n]$, then $\abs{B_n^k(x)} = \abs{P_n(-x)^{h(k)}} (1 - r) \geq \abs{P_n(-x)^{h(k)}} \cdot \frac{2k}{2k+1} \geq 2^{\frac{n(n-1)h(k)}{2}} \cdot \frac{2k}{2k+1}$.
If $x$ is in $[0,1]$, as both $P_n(-x)$ and $-P_n(x)$ are $\leq 0$, we have $\abs{B_n^k(x)} \geq \abs{P_n(-x)^{h(k)}} \geq 2^{\frac{n(n-1)h(k)}{2}}$. Therefore, for all $x \in [0,2^n]$, 
$\abs{B_n^k(x)} \geq \frac 1 2 2^{\frac{n(n-1)h(k)}2}$. The item is concluded by noticing that $B_n^k(x)$ is an even function.

\end{enumerate}

>>>>

\newcommand{\cz}[1]{\textcolor{blue}{Chicheng: #1}}

@@

\section{Construction of polynomials}
Our $\phi$'s in considerations are as follows: TODO: fix the constant in phi so that it has norm at most 1.
\begin{enumerate}
\item $\phi_1(x) = (x_1^{i_1} \ldots x_d^{i_d})_{i_1 + \ldots + i_d \leq \deg, i_1, \ldots, i_d \geq 0} = (x_S)_{S \in N^d, \|S\|_1 \leq \deg}$. This corresponds to the complete symmetric kernel $K_{\deg}$.
\item $\phi_2(x) = ( \sqrt{ {\deg \choose i_1+\ldots+i_d} {i_1+\ldots+i_d \choose i_1 \ldots i_d}}  x_1^{i_1} \ldots x_d^{i_d})_{i_1 + \ldots + i_d \leq \deg, i_1, \ldots, i_d \geq 0}$. This corresponds to the polynomial kernel
$K_{\deg}'(x,x') = (1 + \inner{x, x'})^{\deg}$.
\item $\phi_3(x) = ( \sqrt{ (\frac 1 2)^{i_1+\ldots+i_d} {i_1+\ldots+i_d \choose i_1 \ldots i_d}}  x_1^{i_1} \ldots x_d^{i_d})_{i_1 + \ldots + i_d \leq \deg, i_1, \ldots, i_d \geq 0}$. This corresponds to the kernel $K(x,x') = \frac{1}{1 - \frac 1 2 \inner{x,x'}}$.
\end{enumerate}

@@

\subsection{Construction using rational functions}
\paragraph{The construction.} Define rational function
\[
  Q(x) = \sum_{i=1}^t S_{\ceil{\log \frac 1 \gamma}}^{2t}(\frac{v_i \cdot x}{\gamma}) - (t - \frac 1 2).
\]
where function
\[
S_n^k(x) = \frac{P_n(-x)^{h(k)} - P_n(x)^{h(k)}}{P_n(-x)^{h(k)} + P_n(x)^{h(k)}},
\]
$h(k)$ is the smallest odd integer greater than $\log(2k+1)$, and $P_n(x) = (x - 1) \cdot \prod_{i=1}^n (x - 2^i)^2$.

Now denote by $A_n^k(x) = P_n(x)^{h(k)} - P_n(-x)^{h(k)}$, $B_n^k(x) = -P_n(-x)^{h(k)} - P_n(x)^{h(k)}$.
\cz{Note the difference between this line and KS04 - we multiply by $B_{\ceil{\log \frac 1 \gamma}}^{2t}(\frac{v_i \cdot x}\gamma)$ as opposed to its square. The latter would also work, but it will make the equations more cumbersome.}
Define polynomial $p$ as
\begin{eqnarray*}
  p(x) &:=& \sum_{i=1}^t A_{\ceil{\log \frac 1 \gamma}}^{2t}(\frac{v_i \cdot x}{\gamma}) \cdot \prod_{j \neq i} B_{\ceil{\log \frac 1 \gamma}}^{2t}(\frac{v_i \cdot x}{\gamma}) - (t-\frac12) \prod_{j=1}^t B_{\ceil{\log \frac 1 \gamma}}^{2t}(\frac{v_i \cdot x}{\gamma})
\end{eqnarray*}

It can be seen that
$p(x) = \prod_{i=1}^t B_{\ceil{\log \frac 1 \gamma}}^{2t}(\frac{v_i \cdot x}\gamma) \cdot Q(x)$.

\begin{lemma}[~\cite{klivans2004learning}]
For all $(x,y)$ in $S$ satisfying Definition~\ref{def:int-margin}, we have that $y Q(x) \geq \frac 1 2$.
As a consequence, $y p(x) \geq 2^{\frac{n(n-1)h(2t)t}2 - 2}$.
\end{lemma}
\begin{proof}
We consider two cases regarding the label of the example.
\begin{enumerate}
\item Suppose $y = 1$, then for all $i$ in $[t]$, $v_i \cdot x \in [\gamma, 1]$. This implies that $\frac{v_i \cdot x}{\gamma} \in [1, \frac 1 \gamma]$.
By item~\ref{item:sign-rt} of Fact~\ref{fact:newman}, we have that $D_i(x) \in [1, 1+ \frac 1 {2t}]$.
Hence, $Q(x) \geq t - (t - \frac 1 2) \geq \frac 1 2$.

\item Suppose $y = -1$. First observe that for all $i$ in $[t]$, $\frac{v_i \cdot x}{\gamma} \in [-\frac 1\gamma, \frac 1 \gamma]$. By items~\ref{item:sign-rt},
~\ref{item:sign-lt} and~\ref{item:sign-mid} of Fact~\ref{fact:newman}, we have that $S_{\ceil{\log \frac 1 \gamma}}^{2t}(\frac{v_i \cdot x}{\gamma}) \leq 1+ \frac 1 {2t}$.
In addition, by Definition~\ref{def:int-margin}, there exists an $i_0$ in $[t]$, such that $v_{i_0} \cdot x \leq -\gamma$.
By item~\ref{item:bound-out} of Fact~\ref{fact:cheb}, we have that
$D_{i_0}(x) \in [-1 - \frac 1 {2t}, -1]$.
Hence,
\[ Q(x) = D_{i_0}(x) + \sum_{i \neq i_0} D_i(x) - (t - \frac 1 2)
\leq -1 + (t-1)(1 + \frac 1 {2t}) - (t - \frac 1 2) \leq - \frac 1 2. \]
%\leq -1 + t-1 + \frac{t-1}{2t} - (t - \frac 1 2)
\end{enumerate}
To summarize, we have shown that for all $(x,y)$ in $S$, $y Q(x) \geq \frac 1 2$.
Therefore,
\[ y p(x) = y Q(x) \cdot \prod_{i=1}^t B_{\ceil{\log \frac 1 \gamma}}^{2t}(\frac{v_i \cdot x}\gamma) \geq \frac 1 2 ((1-\frac{1}{4t+1}) \cdot 2^{\frac{n(n-1)h(2t)}2})^t \geq 2^{\frac{n(n-1)h(2t)t}2 - 2}. \]
\end{proof}

\begin{fact}
$S_n^k(\cdot)$ has the following properties:
\begin{enumerate}
\item If $x$ is in $[1, 2^n]$, then $S_n^k(x)$ is in $[1, 1 + \frac 1 k]$. \label{item:sign-rt}
\item If $x$ is in $[-2^{-n}, -1]$, then $S_n^k(x)$ is in $[-1 - \frac 1 k, -1]$. \label{item:sign-lt}
\item If $x$ is in $[-1, 1]$, then $|S_n^k(x)|$ is in $[-1, 1]$. \label{item:sign-mid}
\item For all $x$ in $[-2^{-n}, 2^{-n}]$, $B_n^k(x) \geq (1-\frac{1}{2k+1}) \cdot 2^{\frac{n(n-1)h(k)}2}$. \label{item:sign-denom}
\end{enumerate}
\label{fact:newman}
\end{fact}
\begin{proof}
We prove all the items accordingly.
\begin{enumerate}
\item Observe that $S_n^k(x)$ can be written as:
\[ S_n^k(x) = \frac{1 + (-\frac{P_n(x)}{P_n(-x)})^{h(k)}}{ 1 - (-\frac{P_n(x)}{P_n(-x)})^{h(k)} }.  \]
Denote by $r = (-\frac{P_n(x)}{P_n(-x)})^{h(k)}$. By item~\ref{item:x-large} of Fact~\ref{fact:pn}, $r \in [0, \frac{1}{2k+1}]$. This implies that
$S_n^k(x) = \frac{1+r}{1-r} \in [1, 1 + \frac 1 k]$.

@@

\item Noticing that $B_n^k(x)$ is an even function, without loss of generality, suppose $x \geq 0$.
First, note that if $x \geq 0$, then $\abs{P_n(-x)} \geq \prod_{i=1}^n 2^{2i} = 2^{n(n-1)}$.
If $x$ is in $[1,2^n]$, then $B_n^k(x) = \abs{P_n(-x)^{h(k)}} (1 - r) \geq \abs{P_n(-x)^{h(k)}} \cdot \frac{2k}{2k+1} \geq 2^{n(n-1)h(k)} \cdot \frac{2k}{2k+1}$.
If $x$ is in $[0,1]$, as both $P_n(-x)$ and $-P_n(x)$ are $\leq 0$, we have $B_n^k(x) \geq \abs{P_n(-x)^{h(k)}} \geq 2^{n(n-1)h(k)}$. Therefore, for all $x \in [0,2^n]$,
$\abs{B_n^k(x)} \geq (1-\frac{1}{2k+1}) 2^{n(n-1)h(k)$.
\end{enumerate}
\end{proof}

\paragraph{Norm bound.} We bound the norm of $p$ using the theorem below.
\begin{theorem}
$p(x)$ can be written as $\inner{c, \phi_1(x)}$ for some $c$ such that $\|c\| \leq$.
\end{theorem}
\begin{proof}
Let $n = \ceil{\log \frac 1 \gamma}$.

First, for all $i$ in $[t]$ and $l$ in $\cbr{0,1,\ldots,n}$,
\[ \| (\frac{v_i \cdot x}{\gamma} - 2^l) \|^2 \leq \frac{1}{\gamma^2} + 2^{2l} \leq 2 \cdot 2^{2n}. \]
This implies that
\[ \| P_n(\frac{v_i \cdot x}{\gamma}) \|^2 \leq ((4n+2) \cdot 2^{2n})^{2n+1}. \]
Therefore,
\[ \| P_n(\frac{v_i \cdot x}{\gamma})^{h(k)} \|^2 \leq h(k)^{(2n+1)h(k)} \cdot ((4n+2) \cdot 2^{2n})^{(2n+1)h(k)}. \]
%$A_n^k(x) = P_n(x)^{h(k)} - P_n(-x)^{h(k)}$, $B_n^k(x) = -P_n(-x)^{h(k)} - P_n(x)^{h(k)}$
By the definition of $A_n^k$ and $B_n^k$, we have
\[ \| A_n^k(\frac{v_i \cdot x}{\gamma}) \|^2,  \| B_n^k(\frac{v_i \cdot x}{\gamma}) \|^2 \leq ((8n+4) h(k) \cdot 2^{2n})^{(2n+1)h(k)}. \]

%\sum_{i=1}^t A_{\ceil{\log \frac 1 \gamma}}^{2t}(\frac{v_i \cdot x}{\gamma}) \cdot \prod_{j \neq i} B_{\ceil{\log \frac 1 \gamma}}^{2t}(\frac{v_i \cdot x}{\gamma}) - (t-\frac12) \prod_{j=1}^t B_{\ceil{\log \frac 1 \gamma}}^{2t}(\frac{v_i \cdot x}{\gamma})
Hence, for every $i$ in $[t]$,
\[
\| A_n^{2t}(\frac{v_i \cdot x}{\gamma}) \cdot \prod_{j \neq i} B_n^{2t}(\frac{v_j \cdot x}{\gamma}) \|^2
\leq
t^{(2n+1) h(k) t} \cdot ((8n+4) h(k) \cdot 2^{2n})^{(2n+1)h(k)t}.
\]
Similarly,
\[
\| \prod_{i=1}^t B_n^{2t}(\frac{v_i \cdot x}{\gamma}) \|^2 \leq t^{(2n+1) h(k) t} \cdot ((8n+4) h(k) \cdot 2^{2n})^{(2n+1)h(k)t}.
\]
Therefore,
\[
\| p(x) \|^2 \leq ((t - \frac 1 2)^2 + t) \cdot ((8n+4) h(k) t \cdot 2^{2n})^{(2n+1)h(k)t} \leq ((16n+8) h(k) t \cdot 2^{2n})^{(2n+1)h(k)t}.
\]
where the last inequality uses that $2^{(2n+1)h(k)t} \geq t^2 + \frac 1 4$.
